---
output:
  word_document: default
  html_document: default
---


```{r}
install.packages('gridExtra')
library(creditmodel)
library(gridExtra)
library(tidyverse)
library(ggplot2)
lendingclub
```
# =====================================
# # Univariate Analysis - numeric
# =====================================
# ========================================================
# 1. Create an appropriate plot to visualize the distribution of this variable. (4 marks)
# ========================================================

# =================
# plot 1 - numeric
# =================

# =================
# 
#  This code creates a histogram of the variable "acc_open_past_24mths" in the "lendingclub" dataset. 
# The histogram displays the frequency distribution of the number of trades opened in the past 24 months
#
# The lines of code following the histogram create two additional lines on the plot. The first line 
# adds a red vertical line at the mean value of "acc_open_past_24mths", while the second line adds
# a green vertical line at the median value of the same variable. These lines are added to help 
# visually understand the central tendency of the data. The abline() function is used to add 
# the vertical lines, and the arguments v sets the vertical position of the line, col sets the color of the line,
# and lwd sets the width of the line
# =================
```{r}

acc_open_past_24mths <- lendingclub$acc_open_past_24mths
hist(acc_open_past_24mths, main="Histogram of acc_open_past_24mths", xlab="Number of trades opened in past 24 months", ylab="Frequency", col="blue")

# Add mean and median lines to the histogram
mean_line <- mean(acc_open_past_24mths)
median_line <- median(acc_open_past_24mths)
abline(v = mean_line, col="red", lwd=2)
abline(v = median_line, col="green", lwd=2)
```
# =================
# plot 2 - numeric 
# =================

# This code creates a box plot of the "all_util" variable in the "lendingclub" dataset.
# The box plot is used to visualize the distribution of the data and identify outliers.
# 
# the box plot is drawn vertically (default), with "Balance to credit limit on all trades" as the main title,
# "All trades" as the x-axis label, and "Balance" as the y-axis label. The box color is light blue, 
# and the border is dark blue.

# Create a box plot for all_util
```{r}
# Create a box plot for all_util
boxplot(lendingclub$all_util,
        main = "Balance to credit limit on all trades",
        xlab = "All trades",
        ylab = "Balance",
        col = "lightblue",
        border = "darkblue",
        horizontal = FALSE
)
```
# =================
# plot 3 - numeric
# =================


# This code creates a histogram of the funded_amnt_inv variable in the lendingclub data frame. 
# The main argument sets the title of the histogram, the xlab and ylab arguments label the x and y-axis 
# respectively, and the col argument sets the color of the bars to blue.
# 
# The mean and median of the funded_amnt_inv variable are calculated and plotted as vertical lines 
# on the histogram using the abline() function. The mean line is colored dark magenta and
# the median line is colored green.

```{r}
hist(lendingclub$funded_amnt_inv, main="Histogram of acc_open_past_24mths", xlab="Number of trades opened in past 24 months", ylab="Frequency", col="blue", breaks=6)
mean_line <- mean(lendingclub$funded_amnt_inv)
median_line <- median(lendingclub$funded_amnt_inv)
abline(v = mean_line, col="darkmagenta", lwd=2)
abline(v = median_line, col="green", lwd=2)
```
# ============================================
# Bivariate Analysis
# ============================================

# ========================================
# Create an appropriate plot to visualize the relationship between the two variables.
# =========================================

# =================================
# Both numeric- plot 1 
# =================================

# This code selects two variables from the lendingclub dataset, funded_amnt_inv and installment, 
# and creates a scatterplot with a trend line using ggplot2.

```{r}
# select the variables
selected_vars <- lendingclub %>%
  select(funded_amnt_inv, installment)

# create a scatterplot with trend line
ggplot(selected_vars, aes(x = funded_amnt_inv, y = installment)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Funded Amount", y = "Installment") +
  ggtitle("Relationship between Funded Amount and Installment") +
  theme(plot.title = element_text(hjust = 0.5))
```
# =================================
# Both numeric- plot 2 
# =================================

# This code creates a scatterplot between two variables, total_bal_ex_mort and tot_hi_cred_lim, 
# from the lendingclub dataset.

# The ggplot function is used to create the plot, with selected_vars as the data source.
# The aes function is used to specify the variables to be plotted on the x- and y-axes. 
# The geom_point function adds points to the plot, with color and alpha arguments used to adjust the 
# appearance of the points.

```{r}
# select the variables
selected_vars <- lendingclub %>%
  select(total_bal_ex_mort, tot_hi_cred_lim)

# create a scatterplot
ggplot(selected_vars, aes(x = total_bal_ex_mort, y = tot_hi_cred_lim)) +
  geom_point(color = "#4285F4", alpha = 0.5) +
  labs(x = "Total Balance (excluding mortgage)", y = "Total High Credit/Limit") +
  ggtitle("Relationship between Total Balance and Total High Credit/Limit") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10)) +
  scale_x_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M")) +
  scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M")) +
  theme(axis.line = element_line(color = "black")) +
  geom_smooth()
```
# =================================
# Both numeric- plot 3 
# =================================

# This code creates a scatterplot using the ggplot2 package in R. The data used for the plot is the lendingclub data frame.
# 
# The aes() function is used to define the aesthetics of the plot. The x aesthetic is set to avg_cur_bal, 
# and the y aesthetic is set to mort_acc. These variables represent the average current balance and the 
# number of mortgage accounts, respectively.

```{r}
ggplot(data = lendingclub, aes(x = avg_cur_bal, y = mort_acc)) +
  geom_point(color = "#4285F4", alpha = 0.5) +
  labs(x = "Average Current Balance", y = "Number of Mortgage Accounts") +
  ggtitle("Relationship between Average Current Balance and Mortgage Accounts") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  scale_x_continuous(labels = scales::dollar_format(scale = 1e-3, suffix = "k")) +
  theme(axis.line = element_line(color = "black"),
        axis.line.x = element_line(color = "gray50"),
        axis.line.y = element_line(color = "gray50")) +
  geom_smooth()
```
# =================================
# one numeric and one categorical - plot 1
# =================================

# This code selects the loan_status and all_util variables from the lendingclub dataset and creates
# a new variable all_util_bin by binning the all_util variable into intervals of width 20, 
# with a range from 0 to 200. The loan_status variable is converted to a factor. 
# The code then creates a stacked bar plot with all_util_bin on the x-axis and loan_status on the fill aesthetic. 
# The plot shows the count of loans in each loan_status category for each interval of all_util_bin.
# The legend for loan_status is labeled "Loan Status".

```{r}
# select variables
selected_vars <- lendingclub %>%
  select(loan_status, all_util)

# convert loan_status to a factor
selected_vars$loan_status <- as.factor(selected_vars$loan_status)

# create bins for total credit utilization ratio
selected_vars$all_util_bin <- cut(selected_vars$all_util, 
                                  breaks = seq(0, 200, by = 20), 
                                  labels = seq(0, 180, by = 20))

# create a stacked bar plot
ggplot(selected_vars, aes(x = all_util_bin, fill = loan_status)) +
  geom_bar() +
  labs(x = "Total Credit Utilization Ratio", y = "Count") +
  ggtitle("Loan Status by Total Credit Utilization Ratio") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_discrete(name = "Loan Status")
```
# =================================
# one numeric and one categorical - plot 2 
# =================================

# This code calculates the mean annual income for each loan grade in the lendingclub dataset 
# and creates a bar graph to visualize the results.
# 
# The first line of code uses the group_by() function to group the dataset by loan grade and the summarize()
# function to calculate the mean annual income for each group. The resulting dataframe mean_inc_by_grade 
# has two columns: grade and mean_annual_inc.

```{r}
# calculate mean income for each grade
mean_inc_by_grade <- lendingclub %>%
  group_by(grade) %>%
  summarize(mean_annual_inc = mean(annual_inc))

# create a bar graph
ggplot(mean_inc_by_grade, aes(x = grade, y = mean_annual_inc, fill = grade)) +
  geom_col(alpha = 0.8) +
  labs(x = "Loan Grade", y = "Mean Annual Income", fill = "Loan Grade") +
  ggtitle("Mean Annual Income by Loan Grade") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  theme_minimal()
```

# =================================
# one numeric and one categorical - plot 3
# =================================

# This code selects two variables, dti and loan_status, from the lendingclub dataset. 
# It then converts loan_status to a factor variable. Finally, it creates a scatterplot using ggplot()
# from the ggplot2 package with loan_status on the x-axis, dti on the y-axis, and loan_status 
# as the color variable. The title of the plot is "Relationship between Loan Status and Debt-to-Income Ratio".

```{r}
# select the variables
selected_vars <- lendingclub %>%
  select(dti, loan_status)

# convert loan_status to a factor
selected_vars$loan_status <- as.factor(selected_vars$loan_status)

# create a scatterplot
ggplot(selected_vars, aes(x = loan_status, y = dti, color = loan_status)) +
  geom_point() +
  labs(x = "Loan Status", y = "Debt-to-Income Ratio") +
  ggtitle("Relationship between Loan Status and Debt-to-Income Ratio") +
  theme(plot.title = element_text(hjust = 0.5))
```

